---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# HCC5hmc

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/12379Monty/HCC5hmc.svg?branch=master)](https://travis-ci.com/12379Monty/HCC5hmc)
<!-- badges: end -->

This package contains data on genome-wide mapping of 5-hydroxymethylcytosines
in circulating cell-free DNA as reported in Cai J, Chen L, Zhang Z, et al. Gut
2019;68:2195–2205.

## Installation

The development version from [GitHub](https://github.com/) can be installed with:

``` r
# install.packages("devtools")
devtools::install_github("12379Monty/HCC5hmc")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(HCC5hmc)
## basic example code
```


-----

Please note that the 'HCC5hmc' project is released with a [Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.

<!--

Rscript -e "rmarkdown::render('README.Rmd')"

-->
