---
title: GSE112679 R Data Package
output: github_document   ### will keep md file!!!
bibliography: docs/_bibFiles/bibFile.bib
csl: docs/_bibFiles/cell-numeric.csl
link-citations: true
---

---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


This package contains data from NCBI GEO 
[Series GSE112679](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE112679).
Data methylation profiles were collected to investigate
genome-wide mapping of 5-hydroxymethylcytosines
in circulating cell-free DNA as reported in Cai et al. (2019)
(https://gut.bmj.com/content/68/12/2195).

<!-- [@Cai:2019aa] pandoc citations dont work on github (github_document?) -->



## Data sets  

```{r dataSets}
  library(GSE112679)

  tmp <- data(package='GSE112679')
  knitr::kable(tmp$results[, c("Item", "Title")])

```



## Installation

The development version from [GitHub](https://github.com/) can be installed with:

``` r
# install.packages("devtools")
devtools::install_github("12379Monty/GSE112679")
```

## Example


```{r example}
library(GSE112679)
library(magrittr)

with(sampDesc, table(sampType, trainValGroup, exclude=NULL))

with(sampDesc %>% dplyr::filter(sampType=='blood'), table(outcome, trainValGroup, exclude=NULL))

with(sampDesc %>% dplyr::filter(sampType=='blood'), table(outcome2, trainValGroup, exclude=NULL))

with(sampDesc %>% dplyr::filter(sampType=='blood'), table(outcome3, trainValGroup, exclude=NULL))

```

## Package page

* https://12379monty.github.io/GSE112679/

<!-- DOESNT WORK WITH GITHUB
## References
<div id="refs"></div>
-->

-----

Please note that the 'GSE112679' project is released with a [Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.

<!--

Rscript -e "rmarkdown::render('README.Rmd')"

# nohup Rscript -e "rmarkdown::render('README.Rmd')" > README.log  &


-->
